---
title: "STAT380 Activity 02"
author: "Grace Rovinsky, Min Liu, Jiya Patel"
format: pdf
---

## I. Introduction/Background

## II. Data Exploration
A description of the data that incorporates visualizations, descriptive/incisive statistics, and narrative text describing the data (i.e., EDA)
```{r}
#| echo: false
# add narrative text for each visualization
# Descriptive:
#1. Insurance holder ages per region
ggplot(insurance_data, aes(x = age, y = region)) +
  geom_dotplot(binwidth = 0.2)
#2. Sex of insurance holder vs number of prior claims
ggplot(insurance_data) +
 aes(x = priorclaims, y = sex, fill = smoker) +
 geom_boxplot() +
 scale_fill_hue(direction = 1) +
 theme_minimal()
#3. Insurance holder age vs total charges
ggplot(insurance_data, aes(x=age, y=charges))+
  geom_hex(bins=20, linewidth = 1)

# Inferential: confidence interval/regression line

```

## III. Modeling Methods/Strategies
At least two candidate models, professional looking discussion of the candidate models and the process by which you arrived at them. All decisions should be documented and justified.
- Multiple linear regression
```{r}
# Model 1
insuranceData <- data.frame(insurance_data)
insuranceData <- insuranceData %>%
  mutate(
    smokes = ifelse(smoker == "yes", 1, 0))
View(insuranceData)
insuranceModel1 <- lm(charges ~ age + smokes + priorclaims, data = insuranceData)
summary(insuranceModel1)

# Model 2:

```

## IV. Results
A. Initial Results

B. Interpretations
Interpretations of Model Coefficients and Intercept

C. Model Evaluation
Discussion of the candidate/final model, Evaluation/Validation of the Candidate and Final Model(s): Build as complete of a body of evidence both for and against your candidate/final model.

D. Refinements

## V. Final Recommendations
A paragraph to summarize your recommended model and key findings.

## VI. Author Contributions/Responsibilities

# VII. Code Appendix
```{r}
#| echo: false

## Load necessary packages
library(tidyverse)
library(ggplot2)
library(esquisse)

## Read dataset
insurance_data <- read.csv("~/Downloads/insurance_data_S26.csv")

```

## Once we have final model, email Abby for final check to run on model
## Need to state: How we used the data, variable/model selection process, our metric for model performance
