---
title: "STAT380 Activity 02"
author: "Grace Rovinsky, Min Liu, Jiya Patel"
format: pdf
---

## I. Introduction/Background
The Nittany Valley Insurance Company wants to develop a predictive model for health insurance policy charges based on beneficiary characteristics. Understanding the factors that influece insurance costs enables more accurate pricing, risk assesment, and resource allocation. The objective is to build a linear regression model that predicts total medical charges billed to the policy based on available predictor variables. The analysis follows the modeling cycle: exploratory data analysis, model buildings, evaluation, and refinement to deliver actionable insights to the insurance company.

## II. Data Exploration
A description of the data that incorporates visualizations, descriptive/incisive statistics, and narrative text describing the data (i.e., EDA)
```{r}
#| echo: false
The data contains 1,338 observations with 9 variables: age, sex, bmi, children, smoker, region, charges, month, priorclaims.
```

## III. Modeling Methods/Strategies
At least two candidate models, professional looking discussion of the candidate models and the process by which you arrived at them. All decisions should be documented and justified.

## IV. Results
A. Initial Results

B. Interpretations
Interpretations of Model Coefficients and Intercept

C. Model Evaluation
Discussion of the candidate/final model, Evaluation/Validation of the Candidate and Final Model(s): Build as complete of a body of evidence both for and against your candidate/final model.

D. Refinements

## V. Final Recommendations
A paragraph to summarize your recommended model and key findings.

## VI. Author Contributions/Responsibilities

# VII. Code Appendix
```{r}
#| echo: false
## Read dataset
insurance_data <- read.csv("~/Downloads/insurance_data_S26.csv")
```


1. Have investigated the data as fully as possible to help you build your understanding of the data: A description of the data that incorporates visualizations, descriptive/incisive statistics, and narrative text describing the data (i.e., EDA)
2. The generation of at least two candidate models, professional looking discussion of the candidate models and the process by which you arrived at them. All decisions should be documented and justified.
3. Interpretations of Model Coefficients and Intercept: Discussion of the candidate/final model
4. Evaluation/Validation of the Candidate and Final Model(s): Build as complete of a body of evidence both for and against your candidate/final model.
5. Final recommendation to the Nittany Valley Insurance Company: A paragraph to summarize your recommended model and key findings.
6. Author Contributions/Responsibilities section
7. Code Appendix
8. The report should mix data visualizations (plots, graphs, tables), values of statistics (descriptive, incisive, inferential–as applicable), and narrative text.
